Supporting Ecological Decision Making Using Feature-Selection and Variable Importance

### 2019-09-01 analysis/eval-performance.Rmd

# save as latex table
table1 %>%
  ungroup() %>%
  arrange(RMSE) %>%
  slice(1:15) %>%
  xtable(
    type = "latex",
    caption = "Top 15 results for any task/learner/filter combination, sorted by performance.",
    label = "tab:perf-top-15"
  ) %>%
  print(
    file = "code/98-paper/ieee/performance-top-20.tex",
    include.rownames = TRUE,
    latex.environments = c("center"),
    table.placement = "h!",
    caption.placement = "top"
  )

@@

# save as latex table
table2 %>%
  xtable(
    type = "latex",
    caption = "Best performance of each learner across any task and filter method.",
    label = "tab:best-learner-perf"
  ) %>%
  print(
    file = "code/98-paper/ieee/performance-best-per-learner.tex",
    include.rownames = TRUE,
    latex.environments = c("center"),
    table.placement = "h!",
    caption.placement = "top"
  )

>>>>

# save as latex table
table1 %>%
  ungroup() %>%
  arrange(RMSE) %>%
  slice(1:15) %>%
  xtable(
    type = "latex",
    caption = "Top 15 results for any task/learner/filter combination, sorted by performance.",
    label = "tab:perf-top-15"
  ) %>%
  print(
    file = "code/98-paper/ieee/performance-top-20.tex",
    include.rownames = TRUE,
    latex.environments = c("center"),
    table.placement = "h!",
    caption.placement = "top",
    timestamp = NULL
  )

@@

# save as latex table
table2 %>%
  xtable(
    type = "latex",
    caption = "Best performance of each learner across any task and filter method.",
    label = "tab:best-learner-perf"
  ) %>%
  print(
    file = "code/98-paper/ieee/performance-best-per-learner.tex",
    include.rownames = TRUE,
    latex.environments = c("center"),
    table.placement = "h!",
    caption.placement = "top",
    timestamp = NULL
  )

### 2019-09-01 analysis/feature-importance.Rmd
add feat imp plan

# load drake objects
drake::loadd(
)

@@

```{r }
# create datasets of most imp features for each feature set
data_vi = data.frame(class = 
                    c("VI", "VI", "VI"),
                  rank = 
                    c(1, 3, 3),
                  wavelength = 
                    c(404, 404, 788),
                  feature = c("FeatureA", "FeatureB", "FeatureC")
)

data_nri = data.frame(class = 
                    c("NRI", "NRI", "NRI"),
                  rank = 
                    c(1, 3, 3),
                  wavelength = 
                    c(4050, 470, 600),
                  feature = c("FeatureA", "FeatureB", "FeatureC")
)

# create spectral signature of vegetation 
# PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`
# reflectance is scaled to 0-10 to be able to plot it in the same plot as the filter rankings -> the y-axes need to be the same
# PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm to 1000 nm
# because we order from 1 - 10 with 1 being the best, we have to reverse the scaling of the reflectance values
spectra_sim = hsdar::PROSAIL()
spectra_df = data.frame(reflectance = as.vector(spectra_sim@spectra@spectra_ma),
                wavelength = seq(400, 2500, 1)) %>% 
  dplyr::filter(wavelength < 1000) %>% 
  # scale the reflectance to fit into the range of the y-axis for the filter ranking (the 10 - is to reverse the scale)
  dplyr::mutate(reflectance = 10 - scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10))

# to be able to plot the sec y axis on the plot, we need to supply the scaled R object -> we need to save the object separately
reflectance = as.vector(spectra_sim@spectra@spectra_ma)
reflectance_scaled = scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10)

# here we bind the simulated data with the filter rankings
# Important: Append the feature set class to the simulated data so when merged all together the facetting works (only works if we have class levels for all entries)
# This is also the reason why we cannot create an initial data.frame containing the results of all feature sets
data_vi_merged = bind_rows(spectra_df, data_vi) %>% 
  mutate(class = "VI")

data_nri_merged = bind_rows(spectra_df, data_nri) %>% 
  mutate(class = "NRI")

data_all = bind_rows(data_vi_merged, data_nri_merged)

ggscatter(data_all, x = "wavelength", y = "rank", color = "class", 
          #add = "segments",
          palette = "nejm",
          size = 3,
          facet.by = "class"
) +
  geom_segment(aes(x = wavelength, y = 10, xend = wavelength, yend = rank), color = "grey") + 
  #scale_y_continuous(limits = c(0L, 10L), breaks = scales::pretty_breaks()) +
  scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
  scale_y_reverse(limits = c(10L, 0L), breaks = scales::pretty_breaks(), sec.axis = sec_axis(~ scale(-., center = FALSE, scale = max(., na.rm = TRUE)/-1 ),
                                                                                             labels = c(10, 0.75, 0.55, 0.25, 0),
                                                                                             name = "Reflectance [%]")) +
  labs(y = "Importance", x = "Wavelength [nm]", color = "Feature set") +
  geom_line(aes(x = wavelength, y = reflectance), linetype = "dotted") +
  ggrepel::geom_label_repel(label = data_all$feature, nudge_x = 0.5, nudge_y = 0.5)
```

>>>>

# load drake objects
drake::loadd(fi_permut_vi, fi_permut_nri, fi_permut_hr)

@@

```{r }
# create datasets of most imp features for each feature set
data_vi <- data.frame(
  class =
    c("VI", "VI", "VI"),
  rank =
    c(1, 3, 3),
  wavelength =
    c(404, 404, 788),
  feature = c("FeatureA", "FeatureB", "FeatureC")
)

data_nri <- data.frame(
  class =
    c("NRI", "NRI", "NRI"),
  rank =
    c(1, 3, 3),
  wavelength =
    c(4050, 470, 600),
  feature = c("FeatureA", "FeatureB", "FeatureC")
)

# create spectral signature of vegetation
# PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`
# reflectance is scaled to 0-10 to be able to plot it in the same plot as the filter rankings -> the y-axes need to be the same
# PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm to 1000 nm
# because we order from 1 - 10 with 1 being the best, we have to reverse the scaling of the reflectance values
spectra_sim <- hsdar::PROSAIL()
spectra_df <- data.frame(
  reflectance = as.vector(spectra_sim@spectra@spectra_ma),
  wavelength = seq(400, 2500, 1)
) %>%
  dplyr::filter(wavelength < 1000) %>%
  # scale the reflectance to fit into the range of the y-axis for the filter ranking (the 10 - is to reverse the scale)
  dplyr::mutate(reflectance = 10 - scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE) / 10))

# to be able to plot the sec y axis on the plot, we need to supply the scaled R object -> we need to save the object separately
reflectance <- as.vector(spectra_sim@spectra@spectra_ma)
reflectance_scaled <- scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE) / 10)

# here we bind the simulated data with the filter rankings
# Important: Append the feature set class to the simulated data so when merged all together the facetting works (only works if we have class levels for all entries)
# This is also the reason why we cannot create an initial data.frame containing the results of all feature sets
data_vi_merged <- bind_rows(spectra_df, data_vi) %>%
  mutate(class = "VI")

data_nri_merged <- bind_rows(spectra_df, data_nri) %>%
  mutate(class = "NRI")

data_all <- bind_rows(data_vi_merged, data_nri_merged)

ggscatter(data_all,
  x = "wavelength", y = "rank", color = "class",
  # add = "segments",
  palette = "nejm",
  size = 3,
  facet.by = "class"
) +
  geom_segment(aes(x = wavelength, y = 10, xend = wavelength, yend = rank), color = "grey") +
  # scale_y_continuous(limits = c(0L, 10L), breaks = scales::pretty_breaks()) +
  scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
  scale_y_reverse(limits = c(10L, 0L), breaks = scales::pretty_breaks(), sec.axis = sec_axis(~ scale(-., center = FALSE, scale = max(., na.rm = TRUE) / -1),
    labels = c(10, 0.75, 0.55, 0.25, 0),
    name = "Reflectance [%]"
  )) +
  labs(y = "Importance", x = "Wavelength [nm]", color = "Feature set") +
  geom_line(aes(x = wavelength, y = reflectance), linetype = "dotted") +
  ggrepel::geom_label_repel(label = data_all$feature, nudge_x = 0.5, nudge_y = 0.5)
```

### 2019-09-01 analysis/filter-correlation.Rmd

#source(here::here("code/99-packages.R"))
library("dplyr")
library("tidyr")
library("ggcorrplot")
library("purrr")
library("glue")
library("fs")
R.utils::sourceDirectory("R")

@@

```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
# recode levels so that they are unique
fv_all_list = map2(list(fv_vi_info.gain5$data, fv_vi_info.gain10$data,
  fv_vi_info.gain15$data, fv_vi_info.gain20$data,
  fv_vi_info.gain25$data, fv_vi_info.gain30$data),
  list(5, 10, 15, 20, 25, 30), ~
    dplyr::mutate(.x, method = recode(.x$method, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}"))))

>>>>

# source(here::here("code/99-packages.R"))
library("dplyr")
library("tidyr")
library("ggcorrplot")
library("purrr")
library("glue")
library("fs")
R.utils::sourceDirectory("R")

@@

```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
# recode levels so that they are unique
fv_all_list <- map2(
  list(
    fv_vi_info.gain5$data, fv_vi_info.gain10$data,
    fv_vi_info.gain15$data, fv_vi_info.gain20$data,
    fv_vi_info.gain25$data, fv_vi_info.gain30$data
  ),
  list(5, 10, 15, 20, 25, 30), ~
  dplyr::mutate(.x, method = recode(.x$method, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}")))
)

### 2019-09-01 code/98-paper/ieee/performance-best-per-learner.tex

% latex table generated in R 3.5.2 by xtable 1.8-3 package
% Sun Sep  1 10:28:46 2019
\begin{table}[h!]
\centering
\caption{Best performance of each learner across any task and filter method.} 
\label{tab:best-learner-perf}
\begin{tabular}{rlllr}
  \hline
 & Task & Learner Group & Filter & RMSE \\ 
  \hline
1 & defoliation-all-plots-NRI & RF & Car & 33.96 \\ 
  2 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 33.98 \\ 
  3 & defoliation-all-plots-HR-VI & SVM & Borda & 34.25 \\ 
  4 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
  5 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.56 \\ 
  6 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

% latex table generated in R 3.5.2 by xtable 1.8-3 package
% Sun Sep  1 15:08:58 2019
\begin{table}[h!]
\centering
\caption{Best performance of each learner across any task and filter method.} 
\label{tab:best-learner-perf}
\begin{tabular}{rlllr}
  \hline
 & Task & Learner Group & Filter & RMSE \\ 
  \hline
1 & defoliation-all-plots-HR-NRI-VI & XGBoost & No Filter & 33.95 \\ 
  2 & defoliation-all-plots-NRI & RF & Car & 33.96 \\ 
  3 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 33.98 \\ 
  4 & defoliation-all-plots-HR-VI & SVM & Borda & 34.25 \\ 
  5 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
  6 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.56 \\ 
  7 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
   \hline
\end{tabular}
\end{table}
